---
title: "Data Preparation"
author: "Group 1: Aleksei Luchinsky,  Jingyi   Su, Kimberlyn Brooks,Vibhuti Chandna"
date: "11/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
library(dplyr)
library(xtable)
library(ggplot2)
source("../R/scripts/print_missing.R")
```


```{r}
Dat_orig  <- read.csv("../data/external/Life Expectancy Data.csv")
Dat <- Dat_orig
str(Dat)
```

```{r}
Dat_orig %>% filter( is.na(Population)) %>% select(Country) %>% unique
```
```{r}
Dat_orig %>% group_by(Country) %>% summarise(nmis = sum(is.na(Population))) %>% filter(0 < nmis )
```
```{r}
Dat_orig$Year %>% unique %>% length
```



```{r}
Dat %>% print_missing %>% select(-c(Type, nUnique)) %>% xtable
```


```{r}
year <- 2014
```



Only this year is required

```{r}
Dat <- filter(Dat, Year == year)
miss_2014 <- print_missing(Dat, only_missing = TRUE )
miss_2014
```



```{r}
miss_2014 %>% select(-c(Type, nUnique)) %>% xtable
```


# Population

There are 41 country with missing population. We can try to get this information from other sources, for example https://api.worldbank.org/v2/en/indicator/SP.POP.TOTL?downloadformat=csv


```{r}
data_WB <- 
  read.csv("../data/external/API_SP.POP.TOTL_DS2_en_csv_v2_3158886/API_SP.POP.TOTL_DS2_en_csv_v2_3158886.csv", skip = 3) %>% 
  gather(Year, Population, X1960:X2020) %>% select(Country.Name, Year, Population) %>% 
  transmute(Country = Country.Name, Year = floor(as.numeric(str_sub(Year, 2, 5))), Population.y = Population) 
data_population <- data_WB %>%
  filter(Year == year)
head(data_population)
```

The names of the countries in these tables are not exactly the same, so some dictionary is required:

```{r}
who_ctr <- unique(Dat$Country)
country_dict <- data.frame(WHO = who_ctr, Alt = who_ctr)
country_dict <- rbind( country_dict, data.frame( WHO = "United States of America", Alt = "United States"))
country_dict <- rbind( country_dict, data.frame(WHO = "Bahamas", Alt = "Bahamas, The" ))
country_dict <- rbind( country_dict, data.frame(WHO = "Côte d'Ivoire", Alt = "Cote d'Ivoire"))
country_dict <- rbind( country_dict, data.frame(WHO = "Democratic People's Republic of Korea" , Alt = "Korea, Dem. People's Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Bolivia (Plurinational State of)" , Alt = "Bolivia"))
country_dict <- rbind( country_dict, data.frame(WHO = "Congo" , Alt = "Congo, Dem. Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Democratic Republic of the Congo" , Alt = "Congo, Dem. Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Czechia" , Alt = "Czech Republic"))
country_dict <- rbind( country_dict, data.frame(WHO = "Yemen" , Alt = "Yemen, Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Egypt" , Alt = "Egypt, Arab Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Gambia" , Alt = "Gambia, The"))
country_dict <- rbind( country_dict, data.frame(WHO = "Iran (Islamic Republic of)" , Alt = "Iran, Islamic Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Lao People's Democratic Republic" , Alt = "Lao PDR"))
country_dict <- rbind( country_dict, data.frame(WHO = "Micronesia (Federated States of)" , Alt = "Micronesia, Fed. Sts."))
country_dict <- rbind( country_dict, data.frame(WHO = "Republic of Korea" , Alt = "Korea, Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Republic of Moldova" , Alt = "Moldova"))
country_dict <- rbind( country_dict, data.frame(WHO = "Saint Kitts and Nevis" , Alt = "St. Kitts and Nevis"))
country_dict <- rbind( country_dict, data.frame(WHO = "Saint Lucia" , Alt = "St. Lucia"))
country_dict <- rbind( country_dict, data.frame(WHO = "Saint Vincent and the Grenadines" , Alt = "St. Vincent and the Grenadines"))
country_dict <- rbind( country_dict, data.frame(WHO = "Slovakia" , Alt = "Slovak Republic"))
country_dict <- rbind( country_dict, data.frame(WHO = "United Kingdom of Great Britain and Northern Ireland" , Alt = "United Kingdom"))
country_dict <- rbind( country_dict, data.frame(WHO = "United Republic of Tanzania" , Alt = "Tanzania"))
country_dict <- rbind( country_dict, data.frame(WHO = "Venezuela (Bolivarian Republic of)" , Alt = "Venezuela, RB"))
country_dict <- rbind( country_dict, data.frame(WHO = "Viet Nam" , Alt = "Vietnam"))
country_dict <- rbind(country_dict, data_frame(WHO = "Brunei Darussalam", Alt = "Brunei"))
country_dict <- rbind(country_dict, data_frame(WHO = "Cabo Verde", Alt ="Cape Verde"))
country_dict <- rbind(country_dict, data_frame(WHO =  "Democratic Republic of the Congo", Alt ="Congo, Rep."))
country_dict <- rbind(country_dict, data_frame(WHO =  "Venezuela (Bolivarian Republic of)", Alt ="Venezuela"))
country_dict <- rbind(country_dict, data_frame(WHO =  "Serbia", Alt ="Serbia, FR(Serbia/Montenegro)"))
country_dict <- rbind(country_dict, data_frame(WHO =  "South Sudan", Alt ="Fm Sudan"))
country_dict <- rbind(country_dict, data_frame(WHO =  "Ethiopia", Alt ="Ethiopia(excludes Eritrea)"))
country_dict <- unique(country_dict)
country_dict
```

```{r}
Dat %>% select(Country, Population) %>% merge(country_dict, by.x = "Country", by.y = "WHO" ) %>% 
  merge(data_population, by.x = "Alt", by.y = "Country") -> merged
merged
```

For most of the countries old and new populations agree, but not for all!!!!

```{r}
merged %>% select(Country, Population, Population.y) %>% mutate(rat = Population/Population.y) -> comp_merged
comp_merged
```

```{r}
country = "Afghanistan"
rbind(
  Dat_orig %>% filter(Country == country) %>% select(Year, Population) %>% add_column(type = "WHO"), 
  data_WB %>% filter(Country == country & Year %in% Dat_orig$Year) %>% rename(Population = Population.y) %>%  select(Year, Population) %>% add_column(type = "WB")
) %>% mutate(Population = 1e-6*Population) %>% 
  ggplot(aes(x=Year, y = Population, group = type, color = type)) + geom_line() + geom_point()
  ylab("Population, millions")
  ggsave("../TeX/figures/Afghanistan_population_comparison.pdf")
```


```{r}
MASS::truehist(comp_merged$rat, nbins = 100)
```


Right now I will simply use the new data (this might be not good for Albania, for example)

```{r}
for(c in Dat$Country) {
  new <- merged[ merged$Country == c, ]$Population.y
  if(length(new)>0 && !is.na(new)) {
  Dat[Dat$Country == c,]$Population <- head(new  ,1)
  }
}
```

```{r}
print_missing(Dat)
```

# GDP

GDP info loaded from http://ghdx.healthdata.org/record/ihme-data/gross-domestic-product-gdp-estimates-country-1950-2015

```{r}
GDP <- read.csv("../data/external/IHME_GLOBAL_GDP_ESTIMATES_1950_2015.csv")
GDP <- filter(GDP, Year == year)
GDP <- select(GDP, c(Country, IHME.GDP.estimate..2005.international.dollars. ))
GDP$Country <- trimws(GDP$Country)
GDP$GDP.y <- GDP$IHME.GDP.estimate..2005.international.dollars. %>% trimws %>% str_replace(",","") %>% as.numeric
GDP$GDP.y <- 1e6*GDP$GDP.y
```

```{r}
Dat_orig %>% 
```


```{r}
Dat_orig %>% filter(is.na(GDP)) %>% group_by(Country) %>% summarise(nmis = n()) %>% arrange(nmis) 
```

```{r}
Dat_orig %>% filter(is.na(GDP)) %>% group_by(Country) %>% summarise(nmis = n()) %>% filter(nmis >8) %>% nrow
```

```{r}
colnames(Dat_orig)
```

```{r}
country_
```


```{r}
countries <- unique(Dat_orig$Country)
calc_year_variation <- function(country, iv) {
  dat <- Dat_orig %>% filter(Country == country) %>% .[,iv] 
  if(any(is.na(dat))) {
    return(NA)
  } else {
    return(sd(dat)/mean(dat))
  }
}
i <- 17
print(colnames(Dat_orig)[i])
vars <- sapply(countries, function(c) calc_var(c, i))
vars <- vars[!is.na(vars)]
max(vars)
names(which.max(vars))
```

```{r}
dat <- Dat_orig %>% filter(Country == countries[1]) %>% .[,13]
rat = dat[2:length(dat)]/dat[1:(length(dat)-1)]
sd(rat)/mean(rat)
```

```{r}
par(mfrow=c(1,2))
plot(dat)
plot(rat, ylim=c(0,2))
```


```{r}
Dat_orig %>% filter(Country == "Norway") %>% ggplot(aes(x=Year, y=GDP)) + geom_line()
```


```{r}
colnames(Dat_orig)[12]
```


```{r}
which( colnames(Dat_orig) == "Population" )
```


Comparing GDP data from two sources

```{r}
Dat %>% select(Country, GDP) %>% merge(country_dict, by.x = "Country", by.y = "WHO" ) %>% 
  merge(GDP, by.x = "Alt", by.y = "Country") -> merged
merged
```

Almost no pattern. Will use GDP data set

```{r}
for(c in Dat$Country) {
  new <- merged[ merged$Country == c, ]$GDP.y
  if(length(new)>0 && !is.na(new)) {
    Dat[Dat$Country == c,]$GDP <- head(new  ,1)
  }
}
```



```{r}
print_missing(Dat)
```


# Remove columns

Here are the columns with more then half data missing

```{r}
miss <- sapply(1:ncol(Dat), function(i) sum(is.na(Dat[,i])))
inds <- which(miss > nrow(Dat)/2)
colnames(Dat)[inds]
```


Se can safely remove them

```{r}
if(length(inds)>0) Dat <- Dat[,-inds]
ncol(Dat)
```

```{r}
print_missing(Dat)
```



# aaa

There are small number of rows with NA's

```{r}
nrow(Dat[!complete.cases(Dat),])
```

Removing it for now

```{r}
Dat_final <- Dat[complete.cases(Dat), ]
print_missing(Dat_final)
```



```{r}
write.csv(Dat_final, paste("../data/processed/prepared_data_",year,".csv", sep=""))
```


```{r}
print(year)
model <- lm(Life.expectancy ~ ., data = Dat_final[,-c(1,2)])
summary(model)
```


