---
title: "Data Preparation"
author: "Group 1: Aleksei Luchinsky,  Jingyi   Su, Kimberlyn Brooks,Vibhuti Chandna"
date: "11/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
library(dplyr)
source("../R/scripts/print_missing.R")
```


```{r}
Dat <- read.csv("../data/external/Life Expectancy Data.csv")
str(Dat)
```

Only 2015 year is required

```{r}
Dat <- filter(Dat, Year == 2015)
print_missing(Dat )
```

There is almost no information about Alcohol, Total.expenditure, so excluding the columns

```{r}
Dat <- subset(Dat, select = -c(Alcohol, Total.expenditure))
print_missing(Dat)
```

# Population

There are 41 country with missing population. We can try to get this information from other sources, for example https://api.worldbank.org/v2/en/indicator/SP.POP.TOTL?downloadformat=csv


```{r}
data_population <- 
  read.csv("../data/external/API_SP.POP.TOTL_DS2_en_csv_v2_3158886/API_SP.POP.TOTL_DS2_en_csv_v2_3158886.csv", skip = 3) %>% 
  gather(Year, Population, X1960:X2020) %>% select(Country.Name, Year, Population) %>% 
  transmute(Country = Country.Name, Year = floor(as.numeric(str_sub(Year, 2, 5))), Population.y = Population) %>% 
  filter(Year == 2015)
head(data_population)
```

The names of the countries in these tables are not exactly the same, so some dictionary is required:

```{r}
who_ctr <- unique(Dat$Country)
country_dict <- data.frame(WHO = who_ctr, Alt = who_ctr)
country_dict <- rbind( country_dict, data.frame( WHO = "United States of America", Alt = "United States"))
country_dict <- rbind( country_dict, data.frame(WHO = "Bahamas", Alt = "Bahamas, The" ))
country_dict <- rbind( country_dict, data.frame(WHO = "Côte d'Ivoire", Alt = "Cote d'Ivoire"))
country_dict <- rbind( country_dict, data.frame(WHO = "Democratic People's Republic of Korea" , Alt = "Korea, Dem. People's Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Bolivia (Plurinational State of)" , Alt = "Bolivia"))
country_dict <- rbind( country_dict, data.frame(WHO = "Congo" , Alt = "Congo, Dem. Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Democratic Republic of the Congo" , Alt = "Congo, Dem. Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Czechia" , Alt = "Czech Republic"))
country_dict <- rbind( country_dict, data.frame(WHO = "Yemen" , Alt = "Yemen, Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Egypt" , Alt = "Egypt, Arab Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Gambia" , Alt = "Gambia, The"))
country_dict <- rbind( country_dict, data.frame(WHO = "Iran (Islamic Republic of)" , Alt = "Iran, Islamic Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Lao People's Democratic Republic" , Alt = "Lao PDR"))
country_dict <- rbind( country_dict, data.frame(WHO = "Micronesia (Federated States of)" , Alt = "Micronesia, Fed. Sts."))
country_dict <- rbind( country_dict, data.frame(WHO = "Republic of Korea" , Alt = "Korea, Rep."))
country_dict <- rbind( country_dict, data.frame(WHO = "Republic of Moldova" , Alt = "Moldova"))
country_dict <- rbind( country_dict, data.frame(WHO = "Saint Kitts and Nevis" , Alt = "St. Kitts and Nevis"))
country_dict <- rbind( country_dict, data.frame(WHO = "Saint Lucia" , Alt = "St. Lucia"))
country_dict <- rbind( country_dict, data.frame(WHO = "Saint Vincent and the Grenadines" , Alt = "St. Vincent and the Grenadines"))
country_dict <- rbind( country_dict, data.frame(WHO = "Slovakia" , Alt = "Slovak Republic"))
country_dict <- rbind( country_dict, data.frame(WHO = "United Kingdom of Great Britain and Northern Ireland" , Alt = "United Kingdom"))
country_dict <- rbind( country_dict, data.frame(WHO = "United Republic of Tanzania" , Alt = "Tanzania"))
country_dict <- rbind( country_dict, data.frame(WHO = "Venezuela (Bolivarian Republic of)" , Alt = "Venezuela, RB"))
country_dict <- rbind( country_dict, data.frame(WHO = "Viet Nam" , Alt = "Vietnam"))
country_dict <- rbind(country_dict, data_frame(WHO = "Brunei Darussalam", Alt = "Brunei"))
country_dict <- rbind(country_dict, data_frame(WHO = "Cabo Verde", Alt ="Cape Verde"))
country_dict <- rbind(country_dict, data_frame(WHO =  "Democratic Republic of the Congo", Alt ="Congo, Rep."))
country_dict <- rbind(country_dict, data_frame(WHO =  "Venezuela (Bolivarian Republic of)", Alt ="Venezuela"))
country_dict <- rbind(country_dict, data_frame(WHO =  "Serbia", Alt ="Serbia, FR(Serbia/Montenegro)"))
country_dict <- rbind(country_dict, data_frame(WHO =  "South Sudan", Alt ="Fm Sudan"))
country_dict <- rbind(country_dict, data_frame(WHO =  "Ethiopia", Alt ="Ethiopia(excludes Eritrea)"))
country_dict <- unique(country_dict)
country_dict
```

```{r}
Dat %>% select(Country, Population) %>% merge(country_dict, by.x = "Country", by.y = "WHO" ) %>% 
  merge(data_population, by.x = "Alt", by.y = "Country") -> merged
merged
```
For most of the countries old and new populations agree, but not for all!!!!

```{r}
merged %>% select(Country, Population, Population.y) %>% mutate(rat = Population/Population.y)
```

Right now I will simply use the new data (this might be not good for Albania, for example)

```{r}
for(c in Dat$Country) {
  new <- merged[ merged$Country == c, ]$Population.y
  if(length(new)>0 && !is.na(new)) {
  Dat[Dat$Country == c,]$Population <- head(new  ,1)
  }
}
```

```{r}
print_missing(Dat)
```

# GDP

GDP info loaded from http://ghdx.healthdata.org/record/ihme-data/gross-domestic-product-gdp-estimates-country-1950-2015



```{r}
GDP <- read.csv("../data/external/2015_GDP_WITS-Country-Timeseries.csv")
GDP <- GDP[,c(1,3)]
```

Comparing GDP data from two sources

```{r}
Dat %>% select(Country, GDP) %>% merge(country_dict, by.x = "Country", by.y = "WHO" ) %>% 
  merge(GDP, by.x = "Alt", by.y = "Country.Name") -> merged
merged
```

Almost no pattern. Will use GDP data set



```{r}
for(c in Dat$Country) {
  new <- merged[ merged$Country == c, ]$X2015
  if(length(new)>0 && !is.na(new)) {
    Dat[Dat$Country == c,]$GDP <- head(new  ,1)
  }
}
```

```{r}
print_missing(Dat)
```

There are 25 rows with NA's

```{r}
nrow(Dat[!complete.cases(Dat),])
```

Removing it for now

```{r}
Dat_final <- Dat[complete.cases(Dat), ]
print_missing(Dat_final)
```

```{r}
write.csv(Dat_final, "../data/processed/prepared_data.csv")
```

