---
title: "STAT5020, Project"
author: "Alexey Luchinsky"
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE, input = FALSE}
rm(list=ls())
library(car)
library(StepReg)
library(mice)
source("../scripts/print_missing.R")
```


# Lading and cleaning the data

```{r}
Dat <- read.csv("../data/Life Expectancy Data.csv")
df <- data.frame(Dat)
make_info <- function(Dat, order_by = "missing", decreasing = TRUE) {
  info <- data.frame(
    n=1:ncol(Dat), 
    name = colnames(Dat), 
    type = sapply(1:ncol(Dat), function(i) typeof(Dat[,i])),
    unique = sapply(1:ncol(Dat), function(i) length(unique(Dat[,i]))),
    missing = sapply(1:ncol(Dat), function(i) sum(is.na(Dat[,i])))
  )
  info <- info[ order(info[,order_by], decreasing = decreasing),]
  info
}
list_fields <- function(Dat) {
  make_info(Dat, order_by = "n", decreasing = FALSE)
}
list_missing <- function(Dat) {
  make_info(Dat, order_by = "missing", decreasing = TRUE)
}
list_fields(Dat)
```

Let me first get rid of the string columns

```{r}
Dat2 <- Dat[-c(1,3)] # dropping country
Dat2$developed = ifelse(Dat$Status == "Developed", 1, 0) # factoring status
list_fields(Dat2)
```

Looking for missing values

```{r}
print_missing(Dat2)
```

I will drop all the rows with missing the fields with low number of missing rows:

```{r}
Dat3 <- Dat2
n <- 0
repeat {
  n <- n + 1
  low_mis <- list_missing(Dat3) %>% filter(missing < 30 & missing > 0)
  if(nrow(low_mis)>0 && n < 10) {
    i <- low_mis$n[1]
    Dat3 <- Dat3[!is.na(Dat3[,i]),]
  } else {
    break
  }
}
list(n_removed = nrow(Dat2) - nrow(Dat3))
print_missing(Dat3)
```

7 vars are remaining, I can try to impute with the `mice` package

Using asmple method, looking for insertion with best R2

```{r}
file_name = "../data/impData.rds";
if(! file.exists(file_name)) {
  system.time({
    tmp <- capture.output(impData <- mice(Dat3, method = "sample", m=500, seed = 456))
  })
  r2_list <- sapply(1:impData$m, function(i) {
    dat <- complete(impData, i)
    mod <- lm(Life.expectancy ~ ., data = dat[,-1])
    summary(mod)$r.squared
  })
  saveRDS(list(impData, r2_list),  file_name)
}
impData <- readRDS(file_name)[[1]]
r2_list <- readRDS(file_name)[[2]]
hist(r2_list, main = paste( "R2max = ", round(max(r2_list),3), "at", which.max(r2_list)))
```

The best R2 is slightly less then for original data:

```{r}
model <- lm(Life.expectancy ~ ., data = Dat2)
summary(model)$r.squared
```

but I will work with it

```{r}
Dat_imputed <- complete(impData, which.max(r2_list))
```


# Solving Multicollinearity



```{r}
vif_info <- function(dat, yVar = Life.expectancy) {
  mod <- lm(Life.expectancy ~., data = dat)
  vif_dat <- head( sort(vif(mod), decreasing = TRUE), 3)
  vif_max_name <- names(vif_dat)[1]
  cor_max <- sort( cor(dat)[which(colnames(dat) == vif_max_name),], decreasing = TRUE )[2:4]
  sorted_p <- sort(summary( mod)$coefficients[ c(vif_max_name, names(cor_max)), 4], decreasing = TRUE)
  list(
    vif_dat = vif_dat,
    cor_max = cor_max,
    sorted_p = sorted_p
  )
}
```

```{r}
Dat_vif <- Dat_imputed
model_vif <- lm(Life.expectancy ~ ., data = Dat_vif)
summary(model_vif)$r.squared
```


```{r}
vif_info(Dat_vif)
```

We see that the most problematic var is `infant.deaths`. It is strongly correlated with `under.five.deaths`, so we can remove one

```{r}
Dat_vif <- Dat_vif[, -which(colnames(Dat_vif)=="under.five.deaths")]
model_vif <- lm(Life.expectancy ~ ., data = Dat_vif)
summary(model_vif)$r.squared
```


```{r}
vif_info(Dat_vif)
```

The next problemmatic is thinness.5.9.years. It is strongly correlated with thinness.5.9.years and thinness.5.9.years is more significant, so we will remove thinness.5.9.years

```{r}
Dat_vif <- Dat_vif[, -which(colnames(Dat_vif)=="thinness.5.9.years")]
model_vif <- lm(Life.expectancy ~ ., data = Dat_vif)
summary(model_vif)$r.squared
```

After this step the multicollinearity issue is solved

```{r}
vif_info(Dat_vif)
```

```{r}
summary(model_vif)
```


## SetWise

```{r}
swF <- stepwise(Dat_vif, y="Life.expectancy", selection = "forward")
swF$variate
swF
```

```{r}
swB <- stepwise(Dat_vif, y="Life.expectancy", selection = "backward")
swB$variate
```

```{r}
swBD <- stepwise(Dat_vif, y="Life.expectancy", selection = "bidirection")
swBD$variate
```

```{r}
all.equal( sort(swF$variate), sort(swBD$variate))
```

Both `forward` and `bidirection` methods give the same result, `backward` fails

```{r}
ncol(Dat_vif) - length(swF$variate)
```

6 variables should be removed:

```{r}
to_remove <- which( !(colnames(Dat_vif) %in% swF$variate))
to_remove <- to_remove[-2]
colnames(Dat_vif)[to_remove]
```

```{r}
Dat_sw = Dat_vif[,-to_remove]
model_sw = lm(Life.expectancy ~ ., data = Dat_sw)
summary(model_sw)
```

```{r}
vif(model_sw)
```


You see now that the modle seems to be satisfactory: no ingnificant variables, no multicollinearity problem. It is time now to test it

# Testing model

```{r}
Dat_trans <- Dat_sw
Dat_trans$Adult.Mortality <- (Dat_trans$Adult.Mortality)^1.9
Dat_trans$Hepatitis.B <- sqrt(1+Dat_trans$Hepatitis.B)
model_trans <-  lm(Life.expectancy ~ ., data = Dat_trans)
summary(model_trans)
```

```{r}
aov()
```


```{r}
for(i in 2:ncol(Dat_sw)) {
  MASS::truehist(Dat_sw[,i], main = paste(i, colnames(Dat_sw)[i]))
}
```


```{r}
MASS::truehist( model_sw$residuals)
```




```{r}
for(i in 2:ncol(Dat_trans)) {
  plot(Dat_trans[,i], Dat_trans[,1], main = paste(i, colnames(Dat_trans)[i]))
}
```




```{r}
MASS::truehist(model_sw$residuals)
```


```{r}
ncol(Dat5) - length(swF$variate)
```

5 variables shold be removed:

```{r}
to_remove <- which( !(colnames(Dat5) %in% swF$variate))
to_remove <- to_remove[-2]
colnames(Dat5)[to_remove]
```


```{r}
Dat6 <- Dat5[,-to_remove]
colnames(Dat6)
```

```{r}
model6 <- lm(Life.expectancy ~ ., data = Dat6)
summary(model6)
```

We see that all regressors are significant, R2 is reasonable

# Vriable transformations

```{r}
for(i in 2:ncol(Dat7)) {
  hist(Dat7[,i], breaks = 500, main = colnames(Dat7)[i])
}
```


```{r}
Dat7 <- Dat6
Dat7[,8] = log(1+Dat7[,8])
Dat7[,4] = log(1+Dat7[,4])
Dat7[,9] = log(1+Dat7[,9])
```






```{r}
model7 <- lm(Life.expectancy ~ ., data = Dat7)
summary(model7)
```

```{r}
for(i in 2:ncol(Dat7)) {
  plot(Dat7[,i], model7$residuals, main=paste(i, colnames(Dat7)[i]))
  abline(h=0, col="red")
}
```
```{r}
plot(model7$fitted.values, model7$residuals)
abline(h=0, col="red")
```


```{r}
MASS::truehist(model7$residuals, nbins = 50)
curve( dnorm( x, sd = sd(model7$residuals)), add = TRUE, col = "red")
```

```{r}
shapiro.test(model7$residuals)
```

```{r}
shapiro.test(rnorm(100))
```

