---
title: "Transformation"
author: "Alexey Luchinsky"
date: "11/16/2021"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
library(dplyr)
library(stringr)
```

```{r}
year <- 2014
```

# Preparing Data

```{r}
Dat <- read.csv(paste("../data/processed/prepared_data_",year,".csv", sep=""))
str(Dat)
```

Removing unused components:

```{r}
sum(!complete.cases(Dat))
```


```{r}
Dat <- Dat[,-c(1,2,3)]
str(Dat)
```

The initial model

```{r}
sum <- summary(lm(Life.expectancy ~ ., data = Dat))
r20 <- sum$r.squared
sum
```


```{r}
ncol(Dat)
```

```{r}
transform <- function(x, scale) {
  if(scale == 0) return(log(1 + min(x) + x))
  else return(x^scale)
}
```


```{r}
i0 <- 2
Dat2 <- Dat
```


```{r}
i <- 4
s <- 0
plot(transform(Dat[,i],s), Dat[,i0], xlab = colnames(Dat)[i])
Dat2[,i] <- transform(Dat[,i],s)
colnames(Dat2)[i] <- paste("log_", colnames(Dat)[i],sep="")
```

```{r}
i <- 6
s <- 0
plot(transform(Dat[,i],s), Dat[,i0], xlab = colnames(Dat)[i])
Dat2[,i] <- transform(Dat[,i],s)
colnames(Dat2)[i] <- paste("log_", colnames(Dat)[i],sep="")
```

```{r}
i <- 8
s <- 0
plot(transform(Dat[,i],s), Dat[,i0], xlab = colnames(Dat)[i])
Dat2[,i] <- transform(Dat[,i],s)
colnames(Dat2)[i] <- paste("log_", colnames(Dat)[i],sep="")
```

```{r}
i <- 10
s <- 0
plot(transform(Dat[,i],s), Dat[,i0], xlab = colnames(Dat)[i])
Dat2[,i] <- transform(Dat[,i],s)
colnames(Dat2)[i] <- paste("log_", colnames(Dat)[i],sep="")
```

```{r}
i <- 15
s <- 0
plot(transform(Dat[,i],s), Dat[,i0], xlab = colnames(Dat)[i])
Dat2[,i] <- transform(Dat[,i],s)
colnames(Dat2)[i] <- paste("log_", colnames(Dat)[i],sep="")
```

```{r}
i <- 16
s <- 0
plot(transform(Dat[,i],s), Dat[,i0], xlab = colnames(Dat)[i])
Dat2[,i] <- transform(Dat[,i],s)
colnames(Dat2)[i] <- paste("log_", colnames(Dat)[i],sep="")
```

```{r}
plot(log(Dat[,16]), Dat[,2], xlab = colnames(Dat)[16], ylab = colnames(Dat)[2])
```


```{r}
colnames(Dat2)
```


```{r}
model2 <- lm(Life.expectancy ~ ., data = Dat2)
summary(model2)
```

# Multicollinearity

```{r}
library(car)
vif(model2)
```

```{r}
for(i in 1:ncol(Dat2))
  print(paste(i, colnames(Dat2)[i]))
```


```{r}
Dat3 <- Dat2[,-c(7, 15, 16, 18, 19, 20)]
model3 <- lm(Life.expectancy ~ ., data = Dat3)
summary(model3)
```

```{r}
vif(model3)
```

```{r}
#cbind(1:ncol(Dat2), colnames(Dat2))
as.matrix(colnames(Dat3))
```

```{r}
round(cor(Dat3[,-c(1,2)]) , 2)
```

# Stepwise

```{r}
library(StepReg)
```

```{r}
str(Dat3)
```

```{r}
colnames(Dat3)
```


```{r}
Dat3_ <- Dat3
Dat3_ <- cbind(select(Dat3_, Life.expectancy), select(Dat3_, -Life.expectancy))
Dat3_$Status <- ifelse(Dat3_$Status == "Developing", 0, 1)
sw <- stepwise(Dat3_, "Life.expectancy", selection = "bidirection", select = "adjRsq")
```

```{r}
DatSw <- select(Dat3_, c("Life.expectancy", sw$variate[-1]))
modelSw <- lm(Life.expectancy ~ ., data = DatSw)
summary(modelSw)
```



# Interactions

```{r}
i1 <-  2
i2  <- 4
f <- paste("Life.expectancy ~ ", 
      paste(colnames(DatSw)[-c(1, i1, i2)], collapse  = " + "),
      " + ", colnames(DatSw)[i1], "*", colnames(DatSw[i2]))
mod <- lm(formula = f, data = DatSw)
summary(mod)$adj.r.squared
```

```{r}
# make_int_formula <- function(dat, i1, i2, i0 = 1) {
#   names <- colnames(dat)
#   
#   f <- paste(names[i0], "~", 
#       paste(names[-c(i0, i1, i2)], collapse  = " + "))
#   if(i1 > 0 && i2 > 0) {
#     f <- paste(f, " + ", names[i1], "*", names[i2])
#   }
#   f  
# }

make_int_formula <- function(mod, i1, i2) {
  sum <- summary(mod)
  terms <- as.character(sum$terms)
  vars <- as.vector( 
    sapply( strsplit( terms[[3]], "\\+")[[1]], function(T) str_trim(T))
  )
  vars <- vars[ !grepl("\\*", vars)]
#  print(vars)
  vars_wo <- vars
  if(i1>0 && i2>0) {
    vars_wo <- vars[-c(i1, i2)]
  }
#  print(vars_wo)
  f <- paste(terms[[2]], "~", paste(vars_wo, collapse  = " + "))
  if(i1 > 0 && i2 > 0) {
    f <- paste(f, " + ", vars[i1], "*", vars[i2])
  }
  f  
}
make_int_model <- function(mod, i1, i2) {
#  if(i1 == 0 || i2 == 0) return(NA)
  f <- make_int_formula(mod, i1, i2)
  new_mod <- lm(formula = f, data = mod$model)
  new_mod  
}

make_int <- function(mod, i1, i2) {
  new_mod <- make_int_model(mod, i1, i2)
  sum <- summary(new_mod)
  c(i1, i2, sum$r.squared, sum$adj.r.squared)
}

find_best_int <- function(mod, exclude = c()) {
  xxx <- data.frame(matrix(ncol=4, nrow=0))
  xxx <- rbind(xxx, make_int(mod, 0,0))
  max <- ncol(mod$model) -1
   for(i1 in 1:(max-1)) {
     for(i2 in (i1+1):max) {
       int <- make_int(mod, i1, i2)
       if(any(is.na(int))) next;
       xxx <- rbind(xxx, int)
     }
   }
  colnames(xxx) <- c("i1", "i2", "R2", "adjR2")
  arrange(xxx, desc(adjR2))
}
```

Searching for best interaction combinations

```{r}
xxx <- find_best_int(modelSw)
head(xxx)
```

```{r}
plot(xxx$adjR2, ylab = "adjR2")
```


```{r}
modelInt <- make_int_model(modelSw, 1, 4)
summary(modelInt)
```

Adult.Mortality:HIV.AIDS seems to be most interesting

Selecting only significant vars:

```{r}
coeffs <- summary(modelInt)$coefficients
coeffs[coeffs[,4] < 0.15,]
```


```{r}
final_model <- lm(Life.expectancy ~ 
   Alcohol + log_percentage.expenditure + Status + Total.expenditure + Adult.Mortality*HIV.AIDS
, data <- DatSw)
summary(final_model)
```

This model looks fine

# Removing outliers



```{r}
k <- length(final_model$coefficients)-1
n = nrow(DatSw)

#hatvalues(modelSw) #High leverage if hatvalue > 2*(k+1)/n
leverage <- as.vector( which(hatvalues(final_model) > (2*(k+1)/n))) #
```

```{r}
#rstandard(modelSw) #outlier if abs(rstandard) > 3
#which(abs(rstandard(modelSw)) > 3) # Obs 16 127 is flagged
#rstudent(modelSw) #outlier if abs(rstudent) > 3
outliers <- as.vector(which(abs(rstudent(final_model)) > 3)) # Obs 16 127 is flagged
```



```{r}

difft <- as.vector(which(abs(dffits(final_model)) > (2*sqrt((k+2)/(n-k-2)))))
```

```{r}
crooks <- as.vector(which(cooks.distance(final_model) > 1)) #
```

```{r}
covs1 <- as.vector(which(covratio(final_model) > (1 + 3*(k+1)/n))) #
covs2 <- as.vector(which(covratio(final_model) < (1 - 3*(k+1)/n))) # 
```

```{r}
list(
  leverage = leverage,
  outliers = outliers,
difft = difft, 
crooks = crooks,
covs1 = covs1,
covs2 = covs2
)
```

```{r}
table(c(leverage,
outliers,
difft,
crooks,
covs1,
covs2))

```

```{r}
Dat_cleaned <- DatSw[-127];
model_cleaned <- lm(
  Life.expectancy ~  Alcohol + log_percentage.expenditure + Status + Total.expenditure + Adult.Mortality*HIV.AIDS
, data = Dat_cleaned)
summary(model_cleaned)
```


```{r}
res <- model_cleaned$residuals
SW <- shapiro.test(res)$p.value
MASS::truehist(res, main = paste("SW=", round(SW, 3)))
curve(dnorm(x, sd=sd(res)), add = TRUE, col = "red", lwd = 2)
```



```{r}
bads <- as.numeric(which(abs(res) > 2*sd(res)))
length(bads)/nrow(Dat_cleaned)
```

Our bads are subset of the leverage/outliers list

```{r}
sort(unique(c(leverage,outliers,difft,crooks,covs1,covs2)))
sort(bads)
```


```{r}
yhat <- model_cleaned$fitted.values
plot(yhat, res, xlab = "fitted values", ylab = "residuals")
abline(h=c(0, 2*sd(res), -2*sd(res)), lty = c(1, 2, 3))
points(yhat[bads], res[bads], col="red")
text(yhat[bads]+0.3, res[bads]+0.5, labels = bads, cex = 0.5, col = "red")
```

```{r}
summary(model_cleaned)
```


```{r}
par(mfrow = c(1,2))
y <- model_cleaned$fitted.values
#
x <- model_cleaned$model$Total.expenditure
plot(x, y, xlab="Total.expenditure", ylab="Life.expectancy")
points(x[bads], y[bads], col="red")
text(x[bads]+0.25, y[bads], labels = bads, col="red", cex = 0.7)
#
x <- model_cleaned$model$Adult.Mortality
plot(x, y, xlab="Adult.Mortality", ylab="Life.expectancy")
points(x[bads], y[bads], col="red")
text(x[bads]+30, y[bads], labels = bads, col="red", cex = 0.7)

```

# Final

Here is the model with "bad" points removed

```{r}
model <- lm(
  Life.expectancy ~  Alcohol + log_percentage.expenditure + Status + Total.expenditure + Adult.Mortality*HIV.AIDS
, data = Dat_cleaned[-bads,])
summary(model)
```


```{r}
res = model$residuals
par(mfrow=c(1,2))
SW <- shapiro.test(res)$p.value
MASS::truehist(res, main = paste("SW=", round(SW, 2)))
curve(dnorm(x, sd=sd(res)), add=TRUE, col="red", lwd=2, lty=1)
qqnorm(res)
qqline(res)
```


